Joint Power Allocation and Network Slicing In an End to End O-RAN System

### 2019-10-24 Latex/tex/pr.tex
last version of paper

\author{\IEEEauthorblockN{1\textsuperscript{st} Mojdeh Karbalaee Motalleb}
\IEEEauthorblockA{\textit{Electrical and Computer Engineering} \\
\textit{Tehran University}\\
Tehran, Iran \\
mojdeh.karbalaee@ut.ac.ir}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}

@@

xRAN technology, released in April 2018 as the next generation of RAN, has three fundamental features. The Control plane is decoupled from User plane. Besides, a modular eNB software stack is built to operate on common-off-the-shelf (COTS) hardware. Moreover, open north-bound and south-bound interfaces are introduced\cite{xran}.
\begin{figure}[H]
  \centering
    \includegraphics[scale=0.55]{c4}
  \caption{Network sliced O-RAN system}
  \label{fig:c11}
\end{figure} 
To evolve servicing in 5G, separation of elements of software and hardware of network is deployed in network functions virtualization (NFV) technology. In this technology, the functionality of networks is virtualized and divided into blocks of virtual network function (VNF).
The responsibility of wireless systems of the fifth generation covers wide range types of services. In order to provide the requirements of these services, NS is implemented to virtualize the shared physical network into several logical end-to-end networks. Three different types of NS are introduced in \cite{ns1} contains Core Slicing, RAN Slicing and Core-RAN Slicing. In Core-RAN Slicing, each slice of RAN is mapped to slices of Core. Also, UEs classified into a group of services according to their requirements. In addition, each service is connected to one or more Core-RAN slices based on the resource of slices.
Using cloud-computing in BBU-Pool, the performance of the system is enhanced by virtualizing resources
into virtual machines (VMs).
Each VM has a computing processor in order to be run by the Virtual Network Functions (VNFs) which processed arrival data. Also, VNFs are mapped to physical resources through NS techniques\cite{frdl,luong2018novel,luong2018novel1}.

@@

Where,$\bar{\Omega}_{(f,s)}\in \mathbb{C}^{3}$ and $\Omega_{R_{f,s}}, \Omega_{M_{f,s}}, \Omega_{C_{f,s}}$ indicate the amount of required RAM, Memory and CPU, respectively.
Moreover total amount of required RAM, Memory and CPU of all VNFs of a slice is a three dimension vector which is defined as

@@

The amount of RAM, Memory and CPU is denoted by $\tau_{R_{j}}, \tau_{M_{j}}$and $\tau_{C_{j}} $ for $j^{th}$ DC, respectively .

@@

Where, $\boldsymbol{w} = \{w_R, w_M, w_C\}$ are the weight of RAM, Memory and CPU that is used for Algorithm \ref{alg3}.
In this system model the placement of physical data center resources to VNFs is considered. $y_{s,d}$ is a binary variable indicates whether $d^{th}$ data-center is connected to VNFs of $s^{th}$ slice or not.

@@

Assume the power consumption of baseband processing at each data center that is connected to VNFs of a slice is depicted as
$\phi$. So the total power of system for all active DCs that are connected to slices can be represented as  

@@

\phi_{tot} = \sum_{s=1}^{S}\sum_{d=1}^{D_c}y_{s,d}\phi.

@@

In this paper, the main goal is to simultaneously maximize sum-rate and minimize sum-power which is the total power of all RRHs and the total power consumption of baseband processing at all DCs with the presence of constraints which is written as follow, 

@@

 && \quad \forall \mathfrak{z}, \forall s  \label{c22}

@@

the a physical DC is connected to a VNFs of a slice or not. 
\eqref{c11} and \eqref{c12} indicate that the power of each RRH do not exceed the maximum power and power of each UE is a positive integer value, respectively. Also \eqref{c13} shows that the rate of each UE is more than a threshold. \eqref{c14} and \eqref{c15} depicts that the capacity of fronthaul link is limited and the delay of receiving signal should be less than a threshold, respectively.  

@@

that we have enough physical resource for VNFs of each slice.\newline 
The main optimizaiton problem which is formulated as \eqref{constraints} can be decomposed into two independent optimization problem A and B since the variables can be obtained independently. The problem A is defined as 

@@

& \sum_{s=1}^{S}a_{v,s} \geq 1 &&\quad \forall s

@@

and the problem B is 

@@

 && \quad \forall \mathfrak{z}, \forall s\label{eqomega}

@@

Firstly, we need to obtain $\boldsymbol{A}$ by fixing $\boldsymbol{P}$ in the problem \eqref{constraints1} and updating this parameter at the end of each iteration. Two different method is applied to acquire $\boldsymbol{A}$. 
The details of heuristic algorithm are represented in \textbf{Algorithm} \eqref{alg}.  

@@

\State Sort services according to the number of UEs in it and their requirements in descending order.
\State Sort slices according to the weighted linear combination of number of PRBs, RRHs and VNFs in two layers and the Capacity of their resources in descending order. 

@@

In this part, by assuming that $\boldsymbol{A}$ is fixed, the optimal power of UEs in each service is achieved.

@@

The second subproblem can  be solved using Lagrangian function and iterative algorithm. 
Since, Interference is a function of power of UEs, for simplicity, we assume an upper bound $\bar{I}_{u_{(v,i)}}$ for interference (the worst-case). 
In order to make \eqref{constraints1} as a standard form of convex optimization problem, it is required to change the variable of equations \eqref{cc14} and \eqref{cc15}.
Lagrangian function is written as follow

@@

Where,$\mathfrak{D_s}=\frac{1}{D_{s}^{max}-d_{s_1}-d_{s_2}}+\alpha_s$. Also, $\boldsymbol{\lambda}$, $\boldsymbol{\mu}$, $\boldsymbol{\xi}$ and $\boldsymbol{ \kappa}$ are the matrix of Lagrangian multipliers which have non-zero positive elements. Optimal power is obtained from equation \eqref{lagrang} as follow

@@

p_{u(v,i)}^{*} = \frac{\mathfrak{y}_{u(v,i)}\mathfrak{w}_{u(v,i)}-\mathfrak{x}_{u(v,i)}\mathfrak{z}_{u(v,i)}}{\mathfrak{x}_{u(v,i)}\mathfrak{w}_{u(v,i)} }

@@

In \eqref{firstsub} and \eqref{secondsub} the details of solving each part of subproblem is depicted. 
Firstly, by fixing power of each UE, $\boldsymbol{A}$ is obtained using Algorithm \eqref{alg}. Then by using sub-gradient method, \eqref{secondsub} is solved and iteratively solve these problems until variables converge.

@@

In this subsection, we want to solve  \eqref{constraints2} which is the placement of virtual resources to physical resource in order to minimize the number of using DCs. 
To achieve optimum $\boldsymbol{Y}$ heuristic algorithm and Mosek is applied. The details of heuristic algorithm is written in \textbf{Algorithm} \eqref{alg3}. In this algorithm, firstly we sort slices and DCs according to their sum-weighted of their requirements. Secondly, we start mapping from the most needed slices to the DC with the most physical resources. After mapping DCs to slices, if some slices are not admitted, we start admitting residual slices to more than one DCs. At the end, if DC with the lowest physical resources is free and can served instead of DC with highest physical resource, the slices remapped to new DC with lowest physical resource since it has lowest power consumption.  
if a slice does not admit to a specific DC, it is remain for next placement. In next placement the residual slices, map to more than one DC according to their requirements.

@@

\State  $\{ind_{residual} = s|({\sum_{d=1}^{D_c}y_{s,d}==0})\}$
\State Sort residual amount of DCs same as before in descending order.
\State Sort residual slices same as before in descending order.
\For {$r \gets 1$ to $S_{residual}$}

@@

In this section, Simulation and numerical results for the main problem are depicted.
In Fig. \ref{fig:f1}, the ratio of admitted slices is demonstrated for two different number of DCs with different number of slices (the parameters for simulation listed in table \ref{table:1} and \ref{table:2}). In this simulation, it is assumed that each slice can be served by just one DC and it is not admitted by more than one DC. Proposed method is based on Algorithm \ref{alg3} and optimal method is done by MOSEK toolbox.

@@

When we have two DCs, proposed method and optimal method have approximately same ratio of admitted slices. But by increasing the number of DCs to five, performance of proposed method reduced. Using five DCs, the difference between 
proposed method and optimal method in worst case (44 slices) is about $23$ percentage. 

@@

\begin{figure}[H]

@@

In Fig. \ref{fig:f2}, the normalized resource consumption is depicted due to the number of slices (the parameters for simulation listed in table \ref{table:1} and \ref{table:2}). In this simulation, it is assumed that number of DCs is completely enough to cover all slices. The optimality of placement of DCs to slices is measured. It is shown that how much resources of active DCs are not used. For ten slices, the difference between optimal solution and proposed solution is about $15$ percent. 
\begin{figure}[H]

@@

\bibliographystyle{IEEEtran}
\bibliography{ref}
\end{document}

>>>>

\author{\IEEEauthorblockN{1\textsuperscript{st} Mojdeh Karbalaee Motalleb}
\IEEEauthorblockA{\textit{Electrical and Computer Engineering} \\
\textit{Tehran University}\\
Tehran, Iran \\
mojdeh.karbalaee@ut.ac.ir}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Vahid Shah-Mansouri}
\IEEEauthorblockA{\textit{Electrical and Computer Engineering} \\
\textit{Tehran University}\\
Tehran, Iran \\
vmansouri@ut.ac.ir}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Salar Nouri Naghadeh}
\IEEEauthorblockA{\textit{Electrical and Computer Engineering} \\
\textit{Tehran University}\\
Tehran, Iran \\
salar.nouri@ut.ac.ir}

@@

xRAN technology, released in April 2018 as the next generation of RAN, has three fundamental features. The Control plane is decoupled from User plane. Besides, a modular eNB software stack is built to operate on common-off-the-shelf (COTS) hardware. Moreover, open north-bound and south-bound interfaces are introduced\cite{xran}.
\begin{figure}[H]
  \centering
    \includegraphics[scale=0.55]{c5}
  \caption{Network sliced O-RAN system}
  \label{fig:c11}
\end{figure} 
To evolve servicing in 5G, separation of elements of software and hardware of network is deployed in network functions virtualization (NFV) technology. In this technology, the functionality of networks is virtualized and divided into blocks of virtual network function (VNF).
The responsibility of wireless systems of the fifth generation covers wide range types of services. In order to provide the requirements of these services, NS is implemented to virtualize the shared physical network into several logical end-to-end networks. Three different types of NS are introduced in \cite{ns1} contains Core Slicing, RAN Slicing and Core-RAN Slicing. In Core-RAN Slicing, each slice of RAN is mapped to slices of Core. Also, UEs classified into a group of services according to their requirements. In addition, each service is connected to one or more Core-RAN slices based on the resource of slices.
Using cloud-computing in BBU-Pool, the performance of the system is enhanced by virtualizing resources
into virtual machines (VMs).
Each VM has a computing processor in order to be run by the Virtual Network Functions (VNFs) which processed arrival data. Also, VNFs are mapped to physical resources through NS techniques\cite{frdl,luong2018novel,luong2018novel1}.

@@

Where,$\bar{\Omega}_{(f,s)}\in \mathbb{C}^{3}$ and $\Omega_{R_{f,s}}, \Omega_{M_{f,s}}, \Omega_{C_{f,s}}$ indicate the amount of required RAM, Memory, and CPU, respectively.
Moreover total amount of required RAM, Memory, and CPU of all VNFs of a slice is a three-dimensional vector which is defined as

@@

The amount of RAM, Memory, and CPU is denoted by $\tau_{R_{j}}, \tau_{M_{j}}$and $\tau_{C_{j}} $ for $j^{th}$ DC, respectively.

@@

Where, $\boldsymbol{w} = \{w_R, w_M, w_C\}$ is the weight of RAM, Memory, and CPU that is used for Algorithm \ref{alg3}.
In this system model, the placement of physical DC resources to VNFs is considered. $y_{s,d}$ is a binary variable indicates whether $d^{th}$ DC is connected to VNFs of $s^{th}$ slice or not.

@@

Where, $\sum\limits_{s=1}^{S}\sum\limits_{i=1}^{{R}_s}\bar{p}_{r_{(s,i)}}$ is the total power consumption of all RRHs in all slices. Also, $\sum\limits_{v=1}^{V} \sum\limits_{k=1}^{{U}_v}\mathcal{R}_{u_{(v,k)}} $ is the toal rates of all UEs applied for all types of services. 
Assume the power consumption of baseband processing at each DC $d$ that is connected to VNFs of a slice $s$ is depicted as
$\phi_{s,d}$. So the total power of the system for all active DCs that are connected to slices can be represented as  

@@

\phi_{tot} = \sum_{s=1}^{S}\sum_{d=1}^{D_c}y_{s,d}\phi_{s,d}.

@@

In this paper, the main goal is to maximize sum-rate and minimize sum-power (the total power of all RRHs and the total power consumption of baseband processing at all DCs) simultaneously, with the presence of constraints which is written as follow, 

@@

 && \quad \forall \mathfrak{z}, \forall s; \label{c22}

@@

the physical DC is connected to a VNFs of a slice or not. 
\eqref{c11} and \eqref{c12} indicate that the power of each RRH do not exceed the maximum power, and the power of each UE is a positive integer value, respectively. Also \eqref{c13} shows that the rate of each UE is more than a threshold. \eqref{c14} and \eqref{c15} depicts that the capacity of the fronthaul link is limited, and the delay of receiving signal should be less than a threshold, respectively.  

@@

that we have enough physical resources for VNFs of each slice.\newline 
The main optimizaiton problem which, is formulated as \eqref{constraints}, can be decomposed into two independent optimization problems A and B since the variables can be obtained independently. Problem A is defined as 

@@

& \sum_{s=1}^{S}a_{v,s} \geq 1 &&\quad \forall s.

@@

Problem B is 

@@

 && \quad \forall \mathfrak{z}, \forall s. \label{eqomega}

@@

Firstly, we need to obtain $\boldsymbol{A}$ by fixing $\boldsymbol{P}$ in the problem \eqref{constraints1} and updating this parameter at the end of each iteration. Two different methods are applied to acquire $\boldsymbol{A}$. 
The details of the heuristic algorithm are represented in \textbf{Algorithm} \eqref{alg}.  

@@

\State Sort services according to the number of UEs in it and their requirements in the descending order.
\State Sort slices according to the weighted linear combination of number of PRBs, RRHs and VNFs in two layers and the Capacity of their resources in the descending order. 

@@

In this part, by assuming that $\boldsymbol{A}$ is fixed, the power of UEs in each service is achieved.

@@

The second subproblem can  be solved using the Lagrangian function and iterative algorithm. 
Since, Interference is a function of the power of UEs, for simplicity, we assume an upper bound $\bar{I}_{u_{(v,i)}}$ for interference (the worst-case). 
In order to make \eqref{constraints1} as a standard form of a convex optimization problem, it is required to change the variable of equations \eqref{cc14} and \eqref{cc15}.
The Lagrangian function is written as follow

@@

Where, $\mathfrak{D_s}=\frac{1}{D_{s}^{max}-d_{s_1}-d_{s_2}}+\alpha_s$. Also, $\boldsymbol{\lambda}$, $\boldsymbol{\mu}$, $\boldsymbol{\xi}$, and $\boldsymbol{ \kappa}$ are the matrix of Lagrangian multipliers that have non-zero positive elements. Optimal power is obtained from equation \eqref{lagrang} as follow

@@

p_{u(v,i)}^{*} = [\frac{\mathfrak{y}_{u(v,i)}\mathfrak{w}_{u(v,i)}-\mathfrak{x}_{u(v,i)}\mathfrak{z}_{u(v,i)}}{\mathfrak{x}_{u(v,i)}\mathfrak{w}_{u(v,i)} }]^+

@@

In \eqref{firstsub} and \eqref{secondsub}, the details of solving each part of the subproblem are depicted. 
Firstly, by the fixing power of each UE, $\boldsymbol{A}$ is obtained using Algorithm \eqref{alg}. Then by using the sub-gradient method, \eqref{secondsub} is solved and iteratively solves these problems until variables converge.

@@

In this subsection, we want to solve  \eqref{constraints2}, which is the placement of virtual resources to physical resources in order to minimize the number of using DCs. 
To achieve optimum $\boldsymbol{Y}$ heuristic algorithm is applied. The details of the heuristic algorithm are written in \textbf{Algorithm} \eqref{alg3}. In this algorithm, firstly, we sort slices and DCs according to their sum weighted of their requirements. Secondly, we start mapping from the most needed slices to the DC with the most physical resources. After mapping DCs to slices, if some slices are not admitted, we start admitting remaining slices to more than one DCs. In the end, if DC with the lowest physical resources is free and can be served instead of DC with the highest physical resource, the slices remapped to new DC with the lowest physical resource since it has the lowest power consumption.  
If a slice does not admit to a specific DC, it remains for the next placement. In next placement the remaining slices, map to more than one DC according to their requirements.

@@

\State  $\{ind_{remaining} = s|({\sum_{d=1}^{D_c}y_{s,d}==0})\}$
\State Sort remaining amount of DCs same as before in descending order.
\State Sort remaining slices same as before in descending order.
\For {$r \gets 1$ to $S_{remaining}$}

@@

In this section, simulation and numerical results for the main problem are depicted.
 \caption {Simulation Parameter} \label{table:1a} 
  Noise power & -174dBm\\
  Bandwidth & 120 KHZ \\
 Maxmimun transmit Power & 30dBm \\
  Minimum delay &  30msec \\
  Minimum data rate &  1Mbits/sec \\ [1ex] 
In Fig. \ref{fig:f1a}, energy efficiency is depicted for two different numbers of services with the different mean number of UEs in each service (the parameters for simulation listed in table \ref{table:1a}). The optimal method is $0.016$ bits/J/Hz better than the proposed method for $V = 6$ and $E[U_v] = 10$, and also, $0.011$ bits/J/Hz better than the proposed method for $V = 3$ and, $E[U_v] = 10$. As it is shown, the Energy Efficiency is increased as mean number of UEs rises.
\begin{figure}%[H]
    \includegraphics[width=\linewidth]{figure11}
  \caption{Energy Efficiency vs. Mean Number of UEs in each Service}
  \label{fig:f1a}
In Fig. \ref{fig:f1}, the ratio of admitted slices is demonstrated for two different numbers of DCs with the different number of slices (the parameters for simulation listed in table \ref{table:1} and \ref{table:2}). In this simulation, it is assumed that  just one DC can serve each slice, and it is not admitted by more than one DC. The proposed method is based on Algorithm \ref{alg3}, and optimal method is done by the MOSEK toolbox.

@@

When we have two DCs, the proposed method and optimal method have approximately the same ratio of admitted slices. But by increasing the number of DCs to five, the performance of the proposed method reduced. Using five DCs, the difference between the
proposed method and the optimal method in worst case (44 slices) is about $23$ percentage. 

@@

\begin{figure}%[H]

@@

In Fig. \ref{fig:f2}, the normalized resource consumption is depicted due to the number of slices (the parameters for simulation listed in table \ref{table:1} and \ref{table:2}). In this simulation, it is assumed that the number of DCs is entirely enough to cover all slices. The optimality of the placement of DCs to slices is measured. It is shown that how much resources of active DCs are not used. For ten slices, the difference between the optimal solution and the proposed solution is about $15$ percent. 
\begin{figure}%[H]

@@

\section{Conclusion}
In this paper, joint network slicing and power allocation is considered in an end to end O-RAN system.
It is assumed that UEs are classified based on their service requirements. Also, there are number of slices served the services. Each slice, is consist of PRBs, RRHs, and VNFs that run on VMs. The limited fronthaul capacity is considered for the fiber links between RRHs and BBU-pool.
The target is to maximize sum rate and minimize power consumption and energy cost of data centers simultaneously.
The problem is decomposed into two sub-problems. Each sub-problems are solved separately by a heuristic algorithm. For the sub-problem A, energy efficiency vs. mean number of UEs in each service is depicted. The energy efficiency  exceeded by increasing mean number of UEs in each service. For the sub-problem B, two figures are shown. In the first one, the ratio of admitted slices that connect to just one DC, for different number of slices is denoted. In second figure, the normalized resource consumption of DCs is depicted. In each figure, heuristic algorithm is compared with the optimal method and the difference between them is discussed.

\bibliographystyle{IEEEtran}
\bibliography{ref}
\end{document}
